Viscous Transport in Eroding Porous Media

### 2019-07-31 docs/transport/erosion_transport.tex
increase the font of fig. 4 in transport paper
modify fig. 3(a) in transport paper
working on single body text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{s:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Flow-induced erosion deteriorates and reshapes solid material over a
range of scales.  At small scales, such as in groundwater flow,
fluid-mechanical erosion significantly alters the geometry, and
ultimately changes the process of transport through the media.  This
paper extends a quadrature method first outlined by Barnett, Wu, and
Veerapaneni~\cite{bar-wu-vee2015} to simulate dense packings of eroding
bodies, and then studies coarse-grained variables to characterize
transport in eroded geometries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{s:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  A more
detailed description is in our previous work~\cite{qua-moo2018}.  We
consider flows inside a confined geometry $\Omega$ that contains $M$
eroding bodies with boundary $\gamma_\ell$.  The boundary of the
bounding geometry is $\Gamma$, and the fluid domain is bounded by $\bd
\Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$.  On each
eroding body, $\nn$ is the normal vector pointing into the body, and
$\ss$ is the unit tangent vector pointing in the counterclockwise
direction.  The normal vector on $\Gamma$ is the outward unit normal.
Neglecting inertial forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}\\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass} \\
    \uu = 0, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the bodies} \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity} \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{s:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:stokesDLP} and~\eqref{eqn:deformationDLP}, $\uu$
and its deformation tensor are written as a contour integral in $\RR^2$,
respectively.  However, these layer potentials can be rewritten in terms
of Cauchy integrals and its derivatives in $\CC$.  The first step is to
write the Stokes double-layer potential as the sum of a Laplacian
double-layer potential and its gradient
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Next, since the Laplacian double-layer potential can be written as the
Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$,
the Stokes double-layer potential can be written in terms of a complex
integral and its first derivative.  In addition, the derivative of the
Stokes double-layer potential can be written in terms of a complex
integral and its first two derivatives.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{s:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We have shown that the velocity, shear stress, and vorticity of the
double-layer potential can all be written in terms of Cauchy integrals
and its first two derivatives.  Therefore, we require quadrature methods
to evaluate the expressions in equation~\eqref{eqn:cauchy}.  Since the
integrands are all smooth and periodic, the trapezoid rule has spectral
accuracy.  However, the integrand becomes nearly singular as $x$
approaches $\bd\Omega$, and the quadrature error is arbitrarily large at
a fixed resolution.  There have been a variety of near-singular
integration schemes developed.   We opt to use a method first outlined
by Barnett et al.~\cite{bar-wu-vee2015} which we call the {\em
Barycentric quadrature rule}.  Quadrature rules for only the Cauchy
integral and its first derivative are presented, and we extend this
result to include the second derivative.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{s:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
By using an integral equation formulation, we only have to discretize
the one-dimensional boundary of the domain.  Therefore, we easily
resolve complex geometries, and differentiation can be performed with
spectral accuracy in Fourier space.  We discretize each eroding body
$\gamma_k$ with $N_\iin$ points and discretize the outer wall $\Gamma$
with $N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ is
denoted by $\yy_j^0$, and the $j^{th}$ discretization point on
$\gamma_k$ is denoted by $\yy^k_j$.  The discretization points are
initially distributed evenly in arclength, and this equispacing is
maintained for the entire simulation by using a {\thL}
formulation~\cite{hou-low-she1994}.  In addition, we apply mild
smoothing strategies~\cite{qua-moo2018} to avoid the corners that
erosion inevitably develop.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying the Quadrature}
\label{s:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  By using a fast
algorithm such as the fast multipole method~\cite{gre-rok1987}, this
cost can be reduced to $\mathcal{O}(N)$ operations.  However, the
Barycentric quadrature formulas require several $N$-body calculations
which we would like to avoid.  Instead, we use a hybrid method where the
fast multipole accelerate trapezoid rule is used to compute the velocity
at most of the discretization points, and the non-accelerated
Barycentric quadrature rule is used to locally correct the trapezoid
rule.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Integration}
\label{s:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We reuse the time stepping method outlined in our previous
work~\cite{qua-moo2018}.  Rather than tracking the $(x,y)$ coordinates,
the {\thL} variables are tracked. In addition, a tangential velocity
field is added to each body so that the discretization points remain
equispaced for all time.  A second-order Implicit-Explicit method is
used. In particular, the diffusive term corresponding to the curvature
penalization term is discretized implicitly, and all other terms, which
are non-stiff, are treated explicitly.  By using this time stepping
method in conjunction with the curvature penalization and Gaussian
filter, we stably evolve the eroding bodies.   An in depth description
of the time stepping method for erosion is in Section 3.3 of our
previous work~\cite{qua-moo2018}.

In Section~\ref{s:transport}, we we analyze passive tracers suspended in
an eroded geometry.  Give such a geometry $\Omega$, a tracer $\ss(t)$
initialized at $\ss_0$ is governed by 
\begin{align}
  \label{eqn:trajectories}
  \frac{d\ss}{dt} = \uu(\ss), \quad \ss(0) = \ss_0.
\end{align}
To compute such a trajectory, we apply the standard explicit
fourth-order Runge-Kutta method.  To evaluate $\uu(\ss)$, depending on
the proximity of $\ss$ to $\bd\Omega$, we either use the trapezoid rule
or the Barycentric quadrature rule as described in Section~\ref{s:fmm}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{s:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  For porous media, two of
the most important variables are the porosity and the permeability.
However, additional properties of the flow are revealed by computing
the anomalous diffusion rate, the tortuosity, and the distribution the
gap sizes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{s:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section, we apply our new quadrature method to simulate erosion
of several different configurations of eroding bodies.  The fluid domain
is $\Omega$ whose outer solid walls is a rectangle $[-3,3] \times
[-1,1]$ with rounded off corners.  We only place bodies only in $[-1,1]
\times [-1,1]$.  For all simulations, we discretize $\Gamma = \bd
\Omega$ with $N_\out=1024$ points.  For all but the first example, the
Dirichlet boundary condition on $\Gamma$ is a hagen-poiseuille flow.  As
the bodies erode, the flow rate is adjusted to maintain the physical
constant pressure drop boundary condition.  To maintain this pressure
drop, we compute the pressure at near the inlet and outlet at each time
step as described in our previous work~\cite{qua-moo2018}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a single eroding body close to a solid wall.  To compare our
results with a similar numerical experiment performed by Mitchell and
Spagnolie~\cite{mit-spa2017}, rather a pressure drop boundary condition,
we impose the shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
on the outer wall.  In their experiment, Mitchell and Spagnolie place
the center of the eroding body $1.5$ times the body's radius above the
solid wall.  By using the numerical integration scheme described in
Section~\ref{s:method}, the eroding body can be placed much closer to
the solid wall.  In the following experiments, we consider an eroding
body that is initially a circle of radius $r=0.4$ and is placed near the
no-slip wall at $y=-1$.  We discretize the eroding body with
$N_\iin=256$ discretization points, and the center of the eroding body
is initially chosen so that the distance between the body and the solid
wall are $h/2$, $h/4$, and $h/10$, where $h = 2\pi r/N_\iin$.  For all
three simulations, we use the time step size $\Delta t=10^{-4}$.

With this setup, many of the discretization points on the eroding body
are too close to the outer wall for the trapezoid rule to achieve
sufficient accuracy.  However, by using the Barycentric quadrature rule,
the numerical stable simulations are possible.  In
Figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
several equispaced time steps, and the color of indicates the erosion
rate (magnitude of the shear stress).  Comparing our results to Mitchell
and Spagnolie~\cite{mit-spa2017} (Figure 7(c)), we see similar results.
In particular, between the eroding body and the solid wall the flow is
very small and there is no recirculation, and this is consistent with
three-dimensional results~\cite{cah-feu2002}.  Moreover, for all three
distances, the shear stress is positive \todo[inline]{S-H: Is this
right?  Is the shear stress ever negative?  I might expect so since
there are two stagnation points}, so the erosion is smooth.

where the shear stress is smallest near the zero background flow line
$y=-1$.  Moreover, since the shear stress is everywhere positive, the
erosion rate, which is the magnitude of the shear stress, is a smooth
function.  The result is that the eroding shape does not have any
stagnation points where corners will develop.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Flow-induced erosion deteriorates and reshapes solid material over a
range of scales.  At small scales, such as in groundwater flow,
fluid-mechanical erosion significantly alters the geometry, and
ultimately changes the process of transport through the media.  This
paper extends a quadrature method first outlined by Barnett, Wu, and
Veerapaneni~\cite{bar-wu-vee2015} to simulate dense packings of eroding
bodies, and then studies coarse-grained variables to characterize
transport in eroded geometries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  A more
detailed description is in our previous work~\cite{qua-moo2018}.  We
consider flows inside a confined geometry $\Omega$ that contains $M$
eroding bodies with boundary $\gamma_\ell$.  The boundary of the
bounding geometry is $\Gamma$, and the fluid domain is bounded by $\bd
\Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$.  On each
eroding body, $\nn$ is the normal vector pointing into the body, and
$\ss$ is the unit tangent vector pointing in the counterclockwise
direction.  The normal vector on $\Gamma$ is the outward unit normal.
Neglecting inertial forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}\\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass} \\
    \uu = 0, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the bodies} \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity} \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:stokesDLP} and~\eqref{eqn:deformationDLP}, $\uu$
and its deformation tensor are written as a contour integral in $\RR^2$,
respectively.  However, these layer potentials can be rewritten in terms
of Cauchy integrals and its derivatives in $\CC$.  The first step is to
write the Stokes double-layer potential as the sum of a Laplacian
double-layer potential and its gradient
\begin{equation}
  \label{eqn:Stokes2Laplace}
  \begin{aligned}
    \DDD[\eeta](\xx) &= 
      \frac{1}{2\pi} \int_{\bd\Omega} 
        \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
      \frac{1}{2\pi} \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
      &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
      \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
        \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
  \end{aligned}
\end{equation}
Next, since the Laplacian double-layer potential can be written as the
Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$,
the Stokes double-layer potential can be written in terms of a complex
integral and its first derivative.  In addition, the derivative of the
Stokes double-layer potential can be written in terms of a complex
integral and its first two derivatives.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{sec:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We have shown that the velocity, shear stress, and vorticity of the
double-layer potential can all be written in terms of Cauchy integrals
and its first two derivatives.  Therefore, we require quadrature methods
to evaluate the expressions in equation~\eqref{eqn:cauchy}.  Since the
integrands are all smooth and periodic, the trapezoid rule has spectral
accuracy.  However, the integrand becomes nearly singular as $x$
approaches $\bd\Omega$, and the quadrature error is arbitrarily large at
a fixed resolution.  There have been a variety of near-singular
integration schemes developed.   We opt to use a method first outlined
by Barnett et al.~\cite{bar-wu-vee2015} which we call the {\em
Barycentric quadrature rule}.  Quadrature rules for only the Cauchy
integral and its first derivative are presented, and we extend this
result to include the second derivative.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial Discretization}
\label{sec:spatialDiscretization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
By using an integral equation formulation, we only have to discretize
the one-dimensional boundary of the domain.  Therefore, we easily
resolve complex geometries, and differentiation can be performed with
spectral accuracy in Fourier space.  We discretize each eroding body
$\gamma_k$ with $N_\iin$ points and discretize the outer wall $\Gamma$
with $N_\out$ points.  The $j^{th}$ discretization point on $\Gamma$ is
denoted by $\yy_j^0$, and the $j^{th}$ discretization point on
$\gamma_k$ is denoted by $\yy^k_j$.  The discretization points are
initially distributed evenly in arclength, and this equispacing is
maintained for the entire simulation by using a {\thL}
formulation~\cite{hou-low-she1994}.  In addition, we apply mild
smoothing strategies~\cite{qua-moo2018} to avoid the corners that
erosion inevitably develop.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Efficiently Applying the Quadrature}
\label{sec:fmm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  By using a fast
algorithm such as the fast multipole method~\cite{gre-rok1987}, this
cost can be reduced to $\mathcal{O}(N)$ operations.  However, the
Barycentric quadrature formulas require several $N$-body calculations
which we would like to avoid.  Instead, we use a hybrid method where the
fast multipole accelerate trapezoid rule is used to compute the velocity
at most of the discretization points, and the non-accelerated
Barycentric quadrature rule is used to locally correct the trapezoid
rule.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Integration}
\label{sec:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We reuse the time stepping method outlined in our previous
work~\cite{qua-moo2018}.  Rather than tracking the $(x,y)$ coordinates,
the {\thL} variables are tracked. In addition, a tangential velocity
field is added to each body so that the discretization points remain
equispaced for all time.  A second-order Implicit-Explicit method is
used. In particular, the diffusive term corresponding to the curvature
penalization term is discretized implicitly, and all other terms, which
are non-stiff, are treated explicitly.  By using this time stepping
method in conjunction with the curvature penalization and Gaussian
filter, we stably evolve the eroding bodies.   An in depth description
of the time stepping method for erosion is in Section 3.3 of our
previous work~\cite{qua-moo2018}.

In Section~\ref{sec:transport}, we we analyze passive tracers suspended in
an eroded geometry.  Give such a geometry $\Omega$, a tracer $\ss(t)$
initialized at $\ss_0$ is governed by 
\begin{align}
  \label{eqn:trajectories}
  \frac{d\ss}{dt} = \uu(\ss), \quad \ss(0) = \ss_0.
\end{align}
To compute such a trajectory, we apply the standard explicit
fourth-order Runge-Kutta method.  To evaluate $\uu(\ss)$, depending on
the proximity of $\ss$ to $\bd\Omega$, we either use the trapezoid rule
or the Barycentric quadrature rule as described in Section~\ref{sec:fmm}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  For porous media, two of
the most important variables are the porosity and the permeability.
However, additional properties of the flow are revealed by computing
the anomalous diffusion rate, the tortuosity, and the distribution the
gap sizes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section, we apply our new quadrature method to simulate erosion
of several different configurations of eroding bodies.  The fluid domain
is $\Omega$ whose outer solid walls is a rectangle $[-3,3] \times
[-1,1]$ with rounded off corners.  We only place bodies only in $[-1,1]
\times [-1,1]$.  For all simulations, we discretize $\Gamma = \bd
\Omega$ with $N_\out=1024$ points.  For all but the first example, the
Dirichlet boundary condition on $\Gamma$ is a hagen-poiseuille flow.  As
the bodies erode, the flow rate is adjusted to maintain the physical
constant pressure drop boundary condition.  To maintain this pressure
drop, we compute the pressure at near the inlet and outlet at each time
step as described in our previous work~\cite{qua-moo2018}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a single eroding body close to a solid wall.  To compare our
results with a similar numerical experiment performed by Mitchell and
Spagnolie~\cite{mit-spa2017}, rather a pressure drop boundary condition,
we impose the shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
on the outer wall.  In their experiment, Mitchell and Spagnolie place
the center of the eroding body $1.5$ times the body's radius above the
solid wall.  By using the Barycentric quadrature rule described in
Section~\ref{sec:bary} rather than the trapezoid rule, the eroding body
can be placed much closer to the solid wall without requiring an
unreasonable number of discretization points.  We consider an eroding
body that is initially a circle of radius $r=0.4$ and is placed near the
no-slip wall at $y=-1$.  We discretize the eroding body with
$N_\iin=256$ discretization points, and the center of the eroding body
is initially chosen so that the distance between the body and the solid
wall are $h/2$, $h/4$, and $h/10$, where $h = 2\pi r/N_\iin$.  For all
three simulations, we use the time step size $\Delta t=10^{-4}$.

In Figure~\ref{fig:NearWall}, we superimpose the eroding body's shape at
several equispaced time steps.  For the three initial configurations,
the shear stress is positive for all time, but varies over several
orders of magnitude.  Therefore, we color eroding body's boundary
according to the logarithm of the shear stress.  Comparing our results
to Mitchell and Spagnolie~\cite{mit-spa2017} (Figure 7(c)), we see that
the body shapes are similar.  The differences we can attribute to our
assumption of a two-dimensional flow.  In three dimensions, there is no
recirculation between the solid wall and the spherical
body~\cite{cha-feu2003}.  In contrast, between a two dimensional
circular body and a solid wall, there are regions of recirculation on
the fore and aft sides of the body.  In Figure~\ref{fig:NearWall_vort},
we plot the vorticity at the final configuration, and the where the
vorticity is smallest corresponds to the region where there is a slow
recirculation.

Since the shear stress is always positive, the erosion rate is a smooth
function and does not lead to the development of corners.  However,
since the location where the shear stress is minimized is close to
regions of high shear stress, the rate of erosion changes quickly, and
the result is a body that seems to be developing a corner, although it
is not.  Similar to the three dimensional case~\cite{mit-spa2017}, this
region of high curvature develops in the top half of the eroding body.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r1h_shear}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r0d5h_shear}
\includegraphics[width = 0.38 \textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{\label{fig:NearWall} 
  A single body eroding in a Stokes flow. The color denotes the
  logarithm of the shear stress on the body. Therefore, erosion is
  fastest in the red regions (top half)and slowest in the blue regions
  (lower half).  The body is initialized at three different distances
  from the lower wall: $d=h$ (left), $h/2$ (middle), and $h/10$
  (right).}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r1h_vort}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r0d5h_vort}
\includegraphics[width = 0.33 \textwidth]{./figs/1b_0d4r0d1h_vort}
\caption{\label{fig:NearWall_vort} The vorticity of the fluid with a single body eroding 
at time t=0.1001. The initial distances from body to wall are $h$ (left), $h/2$
  (middle), and $h/10$ (right).  }
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 nearly touching bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies initially placed in $[-1,1] \times
[-1,1]$.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  We start by analyzing the effect of
erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  Initially, 62\% of the geometry is eroded
bodies corresponding to a porosity of $0.38$, and the area fraction
decreases to 0 as the bodies erode.  The trend of the area fraction
resembles that of our previous work~\cite{qua-moo2018} (Figure 10(a)),
but with a larger initial area fraction.  We also compute the flow rate
$U$ required to maintain a constant pressure drop across the channel in
Figure~\ref{fig:Eroding20flowrate}(b).  Again, the trend of $U$
resembles that of our previous work~\cite{qua-moo2018} (Figure 10(b)),
except that the initial flow rate is an order of magnitude smaller
because of the larger initial area fraction.  Starting around normalized
time $0.2$, Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear
which indicates that the flow rate can be written as a power law.  Using
a line of best fit, the flow rate is approximately \begin{align} U
\approx \left(\frac{t}{t_f}\right)^{15/4}, \end{align}
\todo[inline]{update this with whatever S-H gets} which is the dashed
line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/porosity20dense_new}
\todo[inline]{SH: Make the y-limits of this figure go from 0 to 0.7 ...
no need to make it go below 0. Done 07312019}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/flow_rate20dense}
\todo[inline]{SH: In this figure, can you superimpose a line of best fit
from about normalized time 0.2 to normalize time 1.  Doing this crudely,
I got a slope of 15/4}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a): The area fraction of 20
nearly touching eroding bodies versus normalized time. (b): The incoming
flow rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as a power law (dashed line) towards the flow rate
$U=1$ that occurs once all the bodies have eroded.}
\end{figure}
